### 2020-01-24 eold.tex
initial submission

\begin{itemize}
\item {\bf RQ1:}  Does mutation analysis of static analysis tools produce actionable results?  That is, do raw mutation kills serve to distinguish tools from each other, or are all tools similar in terms of mutation detection capability?
\item {\bf RQ2:}  Does our approach provide additional information beyond simply counting findings for the original, un-mutated analyzed code?  Do \emph{ratios} differ between programs, or does the number of mutants killed simply reflect the ``verbosity'' of each tool?
\item {\bf RQ3:}  Do the rankings that raw kills and ratios establish agree with other sources of information about the effectiveness of the evaluated tools?  Can we confirm results from other evaluation methods, or informal opinion?
\item {\bf RQ4:}  Do mutation scores for clean programs differ from those for programs for which tools report findings before mutation?
\item {\bf RQ5:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that
 different tools are good at/bad at?  Can we extend this information to improve tools by addressing identified weaknesses?
\item {\bf RQ6:}  How easy is it identify low-hanging fruit for improving tools without prioritization to remove conceptually redundant mutants?
  \end{itemize}

In particular, we consider {\bf RQ2} to be of critical importance; if the mutant ratios for tools differ, then this is clear evidence that our hypothesis that the tendency of mutants to be faults, and to expect that mutated code will, by a sound and precise tool, be flagged as problematic more often than non-mutated code, holds.  This expectation that (some subset of the) mutants can serve as proxies for real, detectable faults is the core concept of our approach.
{\bf RQ4} addresses a concern briefly mentioned in the introduction:  it is possible that warnings for the original code interfere with our definition of detection.  The ideal case for our approach is when a tool report no findings for un-mutated code, and either reports a finding (kills the mutant) or not (does not kill it) when the mutant is introduced.  Chekam et al. showed that the ``clean program assumption'' for testing, the idea that coverage will be similar for faulty and fixed versions of a program, is a threat to the validity of investigations of the relationship between coverage and fault detection \cite{CleanProgram}, and we want to establish that this is unlikely to be the case for our approach.
{\bf RQ5} and {\bf RQ6} receive only preliminary answers, for the smart contract tools.

We used the Universal Mutator \cite{universalmutator,regexpMut}, which aggressively produces large numbers of mutants, to generate all mutants.  We will make available a repository containing our full raw results plus details on contracts and projects selected, for replication and further analysis, upon acceptance (to avoid breaking blinding).

@@

We analyzed three well-known tools for static analysis of Solidity smart contracts: Slither \cite{slither}, SmartCheck \cite{smartcheck}, and Securify \cite{Securify}.  Slither, based on an SSA-based intermediate language (SlithIR \cite{slither}) is a tool produced by Trail of Bits.  SmartCheck, developed by SmartDec, translates Solidity source directly to an XML-based representation, then uses \emph{XPath} patterns to define problems.  Securify, from SRI Systems Lab at ETH Zurich, works at the bytecode level, first parsing and decompiling contracts, then translating to \emph{semantic facts} in order to look for predefined problems.

@@

We could have used a set of high-transaction contracts, or known-important contracts to validate our approach.  However, we knew that one of our goals in the Solidity experiments was to actually improve a mutation analysis tool, and the developers of the static analysis tools use exactly such benchmarks to validate their tools.  Basing our improvements on mutants of the contracts used for evaluation of proposed detectors would introduce a serious bias in our favor: we would be more likely to produce detectors that would have true positives and few false positives on the benchmark contracts.  We therefore instead selected 100 random contracts for which EtherScan (\url{https://etherscan.io/}) has source code, and used this (quite arbitrary) set of contracts from the actual blockchain to compare tools and identify opportunities for improvement.  The collected contracts had a total of 15,980 non-comment source lines, as measured by {\tt cloc}, with a mean size of 159.8 LOC and a median size of 108 LOC.  The largest single contract had 1,127 lines of code.  Smart contracts are typically smaller than other programs, for various reasons, which explains our use of 100 contracts, compared to a much smaller set of projects for Java and Python.  The Universal Mutator generated 46,769 mutants for the 100 contracts.

@@

Figure \ref{fig:solidityvenn} shows the mutants killed by the Solidity analysis tools.  Tables \ref{tab:scoresolidity} and \ref{tab:cleansolidity} provide numeric details of the results, including the \emph{ratio} for each tool, adjusting its mutation scores by its' general tendency to produce findings.  First, a user examining these results would suspect that Slither and SmartCheck are both useful tools, and should likely both be applied in a high-risk security-sensitive context like smart contract development.  Second, a user might suspect that the large number of findings produced, and smaller number of mutants killed, for Securify, mean that including Securify in the static analysis tool stable is a more difficult decision.  On the one hand, Securify does detect nearly as many mutants it alone can identify as SmartCheck.  The large number of findings, and very bad mutant ratio, however, lead us to suspect that many of these ``detected'' mutants are false positives (or, at least, that the problem is not the one Securify identifies).  Extracting the signal from Securify's noise will be difficult.  We also note that while running Slither and SmartCheck on all 46,769 valid mutants was relatively quick (it took about 6 days sequential compute time for Slither and 3 days for SmartCheck), Securify often required many hours to analyze a mutant, and frequently required a few days to analyze a mutant; the full analysis required over three months of compute time.

For our research questions, {\bf RQ1} is clearly answered in the affirmative.  Figure \ref{fig:solidityvenn} shows that the tools address quite different problems, with all tools reporting far more uniquely detected mutants than mutants in common with other tools.  There are only 18 mutants detected by all tools, all of them involving replacement of {\tt msg.sender} (the caller of a smart contract, which may be another smart contract) with {\tt tx.origin} (the original initiator of a sequence of blockchain calls, a ``human'' account).  Use of {\tt tx.origin} is usuallly a bad idea, can lead to incorrect behavior, and may not be relied on to be meaningful at all in future versions of Ethereum (\url{https://ethereum.stackexchange.com/questions/196/how-do-i-make-my-dapp-serenity-proof}).

{\bf RQ2} is also answered in the affirmative.  Counting findings for un-mutated code might suggest that Securify is the best tool, by a wide margin, but in the context of its near-zero mutant ratio, we must suspect that many of the warnings are false positives.  Slither has the best mutant ratio, but the margin between it and SmartCheck confirms that both tools likely provide value.

For {\bf RQ3}, there are only a few tool comparisons in the literature; this is probably due to the fast-moving nature of the blockchain analysis world; the oldest of these tools' publication dates is 2018.  The most extensive is that of Durieux et al. \cite{durieux2019empirical}.  Slither detected 17\% of known vulnerabilities in their analyis, vs. 11\% for SmartCheck and 9\% for Securify. Slither and SmartCheck were also among the four (out of 9) tools that detected vulnerabilities in the most categories; Securify was not.  The overall recommendation of Durieux et al. was to use a combination of Slither and Mythril \cite{mythril-code} for contract analysis.  Parizi et al. \cite{Parizi} also offer a ranking of tools, and determined that SmartCheck was the most effective, and far more so than Securify; unfortunately, they did not include Slither in their set of evaluated tools.

The Slither paper \cite{slither} also provides an evaluation of all three tools.  Their findings counts differ from ours because of different choices (we threw out merely informational results), but these are unrelated to mutation analysis, in any case.  The evaluation only considered reentrancy faults \cite{SurveyAttacks,FC20} (which are sometimes, but only rarely, introduced by mutants).  For reentrancy, Slither performed best on two real-world large contracts, finding subtle bugs in both, SmartCheck detected the problem in one of the two, and Securify detected neither.  For a dataset of 1,000 contracts, SmartCheck had a high false positive rate (over 70\%) but detected more actual reentrancies (209) than Slither (99) or Securify (6).  On the other hand, Slither's low false positive rate of 11\%  makes its results possibly more useful in practice.  It seems safe to say that

For {\bf RQ4}, on the changes seen when restricting analysis to clean contracts, Slither did slightly better at detecting mutants when the original contract was clean for Slither, and the other two tools did somewhat worse on contracts for which they reported no findings.  For the three contracts clean for all tools, Slither performed almost exactly as it did over contracts in general, and the other tools performed worse, by about the same margin as they did for their own clean contracts.  For our approach, we only need a week version of the ``clean program assumption'':  the threat is that kills may be under-reported for non-clean programs, due to interference with findings for the original code.  It is not a problem if mutation scores are \emph{worse} for programs where a tool reports no findings for the un-mutated code.  We therefore, for smart contracts, find no threat to our approach arising from the presence of findings on un-mutated code.  We speculate that ``clean'' results for some tools result from contracts where the tool has trouble with the contract code, but does not actually crash; Slither may do better on clean code because it has fewer such failures, and clean contracts are probably generally simpler and easier to analyze.

\subsubsection{Improving Slither {\bf (RQ5 and RQ6)}}

\begin{figure}
  {\scriptsize
\noindent \begin{code}
 0x598ab825d607ace3b00d8714c0a141c7ae2e6822\_Vault.mutant.275.sol:
 if (!p.recipient.send(p.amount)) \{  // Make the payment
 ==>          if (true) \{  // Make the payment
 if (true) \{  // Make the payment
      \end{code}
      }
      \caption{Mutant showing {\bf Boolean constant misuse}.}
      \label{fig:boolean}
    \end{figure}

    \begin{figure}
      {\scriptsize
\begin{code}
 0x968815CD73647C3af02a740a2438D6f8219e7534\_TTPresale.mutant.311.sol:
 require(nextDiscountTTMTokenId6 >= 361 \&\& nextDiscountTTMTokenId6 <= 391);
 ==>  ...361...==>...0...
 require(nextDiscountTTMTokenId6 >= 0 \&\& nextDiscountTTMTokenId6 <= 391);
      \end{code}
      }
      \caption{Mutant showing {\bf Type-based tautologies}.}
      \label{fig:tautology}      
\end{figure}

\begin{figure}
  {\scriptsize
\begin{code}
 0x534ccee849a688581d1b0c65e7ff317ed10c5ed3\_NametagToken.mutant.480.sol:
 byte char = byte(bytes32(uint(x) * 2 ** (8 * j)));
 ==>  ...*...==>.../...
 byte char = byte(bytes32(uint(x) * 2 ** (8 / j)));
      \end{code}
      }
      \caption{Mutant showing {\bf Loss of precision}.}
      \label{fig:divmul}            
      \end{figure}

@@

@@

\item {\bf Loss of precision:}  Solidity only supports integer types.  This means that performing division before multiplication can introduce rounding that is not present when the multiplication is performed first.  This is a fairly important problem, given the frequency with which Solidity code performs financial calculations where maximum precision is desired.  SmartCheck provides a detector for such precision losses \url{https://github.com/smartdec/smartcheck/blob/master/rule\_descriptions/SOLIDITY\_DIV\_MUL/description\_en.html}, which enabled it to detect mutants such as the one shown in Figure\ref{fig:divmul}.
\end{enumerate}

@@

Examining the first 100 mutants in the unprioritized lists for SmartCheck and Securify, ordered by contract ID and mutant number (roughly source line mutated) we were unable to identify \emph{any} obviously interesting mutants, suggesting that for {\bf RQ6} it is indeed hard to use mutation analysis results without prioritization. A large majority of the mutants we inspected involved either the missing
{\tt return} problem noted in the introduction, or replacing {\tt msg.sender} with {\tt tx.origin}; Slither, of course, has a detector for misuses of {\tt tx.origin}, and we believe (but are not sure) that almost all of the differences with respect to such changes are due to intentional behavior. SmartCheck and Securify tend to identify most (though not all) uses of {\tt tx.origin} as incorrect, while Slither has a more selective rule, intended to reduce false positives.
It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying good candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ5} and {\bf RQ6}.

@@

70,825 mutants

@@

\begin{table}
  \begin{tabular}{l|r|r|r|r|r}
    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & Mutant \\
    Tool & Mean & Median & Mean & Median & Ratio\\
    \hline
    \hline
    SpotBugs & 0.47 & 0.00 & 0.37 & 0.21 & 0.7847 \\
    \hline
    PMD & 1.27 & 0.00 & 0.07 & 0.07 & 0.0562 \\
    \hline
    Infer & 0.07 & 0.00 & 0.00 & 0.00 &  0.0002 \\
    \hline
  \end{tabular}
  \caption{Java tools: number of findings and mutation scores over all projects.}
  \label{tab:scorejava}
\end{table}

\begin{table}
  \begin{tabular}{l|r|r|r}
    & & \multicolumn{2}{|c|}{} \\
    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} \\
    Tool & Projects & Mean & Median \\
    \hline
    \hline
    SpotBugs & 3 & 0.41 & 0.17 \\
    \hline
    PMD & 0 & N/A & N/A \\
    \hline
    Infer & 6 & 0.00 & 0.00 \\
    \hline
  \end{tabular}
  \caption{Java tools: clean project counts and mutation scores.  No projects were clean for all tools, due to PMD producing findings for all projects.}
  \label{tab:cleanjava}
\end{table}

Figure \ref{fig:javavenn} shows the mutants killed by the Java analysis tools, and Tables \ref{tab:scorejava} and \ref{tab:cleanjava} provide numeric results for projects and clean projects, respectively.

In terms of {\bf RQ1}, the raw results are clear:  SpotBugs finds vastly more uniquely killed mutants than the other tools, but PMD contributes a large number of its own unique detections, and clearly offers some independent utility.  Infer on the other hand, is only able to detect two mutants, but these are unique.

{\bf RQ2} is also answered in the affirmative.  While the raw kills for SpotBugs are impressive, it also produced fewer findings for clean code, giving it an excellent mutant ratio, making its dominance even more clear.  PMD's performance was acceptable, and better than any of the Solidity tools (note that comparing ratios across languages is potentially invalid, but not obviously so).  Infer is confirmed as nearly useless, again.

\begin{figure}
  {\scriptsize
  \begin{itemize}
  \item \url{https://stackoverflow.com/questions/4297014/what-are-the-differences-between-pmd-and-findbugs}
  \item \url{https://www.sw-engineering-candies.com/blog-1/comparison-of-findbugs-pmd-and-checkstyle}
  \item \url{https://www.reddit.com/r/java/comments/3i7w6n/checkstyle_vs_pmd_vs_findbugs_for_dummies_why/}
  \end{itemize}
  }
\caption{Blog posts and forum threads discussing Java static analysis tools.}
\label{fig:blog}
\end{figure}

For {\bf RQ3}, to our knowledge there is no academic comparison of these tools.  However, the user postings listed in Figure \ref{fig:blog}, plus personal communications with security analysts who use these tools (citation removed for now, due to potential to compromise blinding) supported some basic conclusions.  SpotBugs/FindBugs is indeed the best tool for finding bugs; PMD focuses more on stylistic issues and has a weaker semantic model.  Mutants are much more likely to be bugs than stylistic issues (they alter semantics, not code style, in almost all cases: deletion or constant/operator changes obviously have no style impact).  Infer is closer to a model checker focusing on resource leaks than a truly general-purpose tool, arguably.

For {\bf RQ4} there were very few clean projects, but the mutation scores for Infer were unchanged, and for SpotBugs the mean mutation score improved and median mutation score was worse.  Again, we see no evidence that non-clean code is a source of significant degradation on mutation detection.

@@

167,511 mutants

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{python.png}
  \caption{Mutants killed by Python static analysis tools.}
  \label{fig:pythonvenn}
\end{figure}

\begin{table}
  \begin{tabular}{l|r|r|r|r|r}
    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & Mutant \\
    Tool & Mean & Median & Mean & Median & Ratio\\
    \hline
    \hline
    Pylint & 2.24 & 1.00 & 0.22 & 0.20 & 0.097 \\
    \hline
    pyflakes & 2.4 & 1.00 & 0.03 & 0.00 & 0.012 \\
    \hline
    PyChecker & 0.56& 0.00 & 0.10 & 0.00 &  0.183 \\
    \hline
  \end{tabular}
  \caption{Python tools: number of findings and mutation scores over all projects.}
  \label{tab:scorepython}
\end{table}

\begin{table}
  \begin{tabular}{l|r|r|r}
    & & \multicolumn{2}{|c|}{} \\
    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} \\
    Tool & Projects & Mean & Median \\
    \hline
    \hline
    Pylint & 0 & N/A & N/A \\
    \hline
    pyflakes & 12 & 0.02 & 0.00 \\
    \hline
    PyChecker & 21& 0.01 & 0.00 \\
    \hline
  \end{tabular}
  \caption{Python tools: clean project counts and mutation scores.  No projects were clean for all tools, due to Pylint producing findings for all projects.}
  \label{tab:cleanpython}
\end{table}

Figure \ref{fig:pythonvenn} shows the mutants killed by the Java analysis tools, and Tables \ref{tab:scorepython} and \ref{tab:cleanpython} provide numeric results for projects and clean projects, respectively.

For {\bf RQ1}, Pylint and pyflakes both killed more than 15,000 mutants killed by no other tool; Pylint killed more mutants, but the relative difference was much smalller than in our Solidity or Java analysis.  It is clear both tools are well worth using, assuming these are not false positives.  Pyflakes' mutation score is lower than unique kills might suggest because it has almost no overlap with the other two tools.  It is doing something quite different than they are, apparently.

The mutant ratios {\bf RQ2} suggest that PyChecker might be better than pyflakes, in fact, and better than Pylint, because while it killed fewer mutants, it also produced fewer findings by far, overall.  PyChecker would be a good choice for quick and dirty analysis of code on the fly, perhaps, in that it had excellent mutation scores for some programs (hence a high mean) but is relatively ``quiet'' compared to the other tools.  However, its very low median mutation score is troubling; it may simply not work for some code.  A combination of pyflakes and Pylint seems safer, though adding the relatively low-findings PyChecker seems likely to have little cost.  For Python, we see that all of our measures are required to get a good picture of the tools.

\begin{figure}
  {\scriptsize
  \begin{itemize}
  \item \url{https://stackoverflow.com/questions/1428872/pylint-pychecker-or-pyflakes}
  \item \url{https://www.reddit.com/r/Python/comments/ii3gm/experience_with_pylint_pychecker_pyflakes/}
  \item \url{https://www.slant.co/versus/12630/12631/~pylint_vs_pyflakes}
  \item \url{https://news.ycombinator.com/item?id=12748885}
  \item \url{https://doughellmann.com/blog/2008/03/01/static-code-analizers-for-python/}
  \end{itemize}
  }
\caption{Blog posts and forum threads discussing Python static analysis tools.}
\label{fig:blogpython}
\end{figure}

For {\bf RQ3}, there were again no academic comparisons we could find.  However, opinions on the web were quite common (see Figure \ref{fig:blogpython}, which lists ones we examined).  It is hard to summarize the overall opinion here, since it ranges considerably.  There is probably general agreement that PyChecker is old and maybe less useful, but also terse and sometimes helpful.  Pylint is the most recommended tool, and the general complaint that it is too picky was mitigated in our results by turning off clearly informative-only findings; users not configuring Pylint will probably see even more killed mutants, but likely a worse mutant ratio because mutants tend to introduce non-stylistic problems.  Pyflakes is also well liked, and supposedly less verbose than Pylint; we did not see this in our results, but that may be because we configured Pylint, and pyflakes aims to focus on ``actual bugs'' so it is equally verbose for mutants, which are likely to be actual bugs.

For {\bf RQ4} there were no clean projects for the known-to-be-aggressive Pylint.  Performance for the other two tools was worse on projects for which they were clean, again failing to support the possibility that our definition of killing a mutant under-reports detection for non-clean code.

@@

The primary threat to validity in terms of generalization is that we only examined nine static analysis tools, and our analyses were restricted to 100 smart contracts, 15 Java projects, and 25 Python projects.   Because it is hard to identify a ground truth to compare with (the motivation for our approach), we cannot be certain that our rankings of tools are correct even for these tools and this code.  However, where there are existing discussions of the tools, our results seem to agree with these, but add substantial detail.

A general threat to our approach is that mutants may differ substantially from ``real'' faults, in some way.  This is certainly true, in a sense \cite{GopinathMutants}, but for static analysis purposes, the risk is that some mutation operators align with patterns a particular tool identifies, biasing the evaluation in favor of that tool.  Such faults may be dis-proportionately present in mutants vs. real code.  However, we consider this unlikely.  The vast majority of applied mutation operations for all of our experiments were highly generic, and do not plausibly represent a pattern in which some tool might specialize.  Statement deletions leave no ``trace'' for a tool to match against, but only an omission.  Changing arithmetic and comparison operators and numeric constants (incrementing, decrementing, or changing to 0 or 1) account for most of the non-statement-deletion mutants.  We cannot see how any tool could have a (useful) detection rule that looks for these kinds of changes (a tool obviously would not flag all instances of 0 and 1 constants, or addition).  Finally, many mutants (though many fewer than in these categories) add {\tt break} or {\tt continue} statements.  Again, we do not see how a tool could take advantage of this without either genuinely understanding likely-bad control flow patterns (in which case, we think it ``deserves'' to kill more mutants) or flagging so many instances of these it would have a bad mutant ratio.

>>>>

\begin{itemize}
\item {\bf RQ1:}  Does our approach produce actionable results?  That is, do raw mutation kills serve to distinguish tools from each other, or are all tools similar in terms of mutation detection capability?
\item {\bf RQ2:}  Does our approach provide additional information beyond simply counting findings for the original, un-mutated analyzed code?  Do \emph{ratios} differ between programs, or does the number of mutants killed simply reflect the ``verbosity'' of each tool?
\item {\bf RQ3:}  Do the rankings that raw kills and ratios establish agree with other sources of information about the effectiveness of the evaluated tools?  Can we confirm results from other evaluation methods, or informal opinion?
\item {\bf RQ4:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that different tools are good at/bad at?  Can we extend this information to improve tools by addressing identified weaknesses?
  \end{itemize}

In particular, we consider {\bf RQ2} to be of critical importance; if the mutant ratios for tools differ, then this is clear evidence that our hypothesis that the tendency of mutants to be faults, and to expect that mutated code will, by a sound and precise tool, be flagged as problematic more often than non-mutated code, holds.  This expectation that (some subset of the) mutants can serve as proxies for real, detectable faults is the core concept of our approach.

@@

We analyzed three well-known tools for static analysis of Solidity smart contracts: Slither \cite{slither}, SmartCheck \cite{smartcheck}, and Securify \cite{Securify}.

@@

We could have used a set of high-transaction contracts, or known-important contracts to validate our approach.  However, we knew that one of our goals in the Solidity experiments was to actually improve a mutation analysis tool, and the developers of the static analysis tools use exactly such benchmarks to validate their tools.  Basing our improvements on mutants of the contracts used for evaluation of proposed detectors would introduce a serious bias in our favor: we would be more likely to produce detectors that would have true positives and few false positives on the benchmark contracts.  We therefore instead selected 100 random contracts for which EtherScan (\url{https://etherscan.io/}) has source code, and used this (quite arbitrary) set of contracts from the actual blockchain to compare tools and identify opportunities for improvement.

@@

Figure \ref{fig:solidityvenn} shows the mutants killed by the Solidity analysis tools.  Tables \ref{tab:scoresolidity} and \ref{tab:cleansolidity} provide numeric details of the results, including the \emph{ratio} for each tool, adjusting its mutation scores by its' general tendency to produce findings.  First, a user examining these results would suspect that Slither and SmartCheck are both useful tools, and should likely both be applied in a high-risk security-sensitive context like smart contract development.  Second, a user might suspect that the large number of findings produced, and smaller number of mutants killed, for Securify, mean that including Securify in the static analysis tool stable is a more difficult decision.  On the one hand, Securify does detect nearly as many mutants it alone can identify as SmartCheck.  The large number of findings, and very bad mutant ratio, however, lead us to suspect that many of these ``detected'' mutants are false positives (or, at least, that the problem is not the one Securify identifies).  Extracting the signal from Securify's noise will be difficult.  We also note that while running Slither and SmartCheck on all 46,769 valid mutants was relatively quick (it took about 8 days sequential compute time for Slither and 3 days for SmartCheck), Securify often required many hours to analyze a mutant, and frequently required a few days to analyze a mutant; the full analysis required over three months of compute time.

We first compare the two tools that performed best.  Slither detected 2.37 mean issues over the 100 contracts, and found 39 contracts free of all issues.  Slither had a mean mutation score of 0.09 over all contracts, and 0.11 over clean (according to Slither) contracts.  SmartCheck detected 1.89 mean issues over the 100 contracts, but only found 27 contracts free of all issues.  The mean issue counts for the tools were not statistically significantly different ($p=0.34$).  SmartCheck had a mean mutation score of 0.05 over all contracts, and 0.03 over clean (according to SmartCheck) contracts.  The ratio between mean issues (Slither / SmartCheck) is 1.25, and the ratio of numbers of clean contracts is similar (1.44).  The mutation score ratio, however, is 1.8, suggesting that Slither is not out-scoring SmartCheck \emph{only} because it produces more false positives.  Slither detects 3,520 mutants not detected by SmartCheck, while SmartCheck only detects 1,629 mutants not detected by Slither (of these, 1,256 were also not detected by Securify) .  Over the 7 contracts that both tools mark as clean, Slither has a mean mutation score of 0.08, and SmartCheck has a mean mutation score of 0.03; the difference, even with only 7 contracts, is statistically significant by paired Wilcoxon test, at $p=0.03$.

Securify was an outlier, compared to the other tools.  First, while running Slither and SmartCheck on all 46,769 valid mutants was relatively quick, Securify often required hours to analyze a contract; the full analysis required over a month.  Second, Securify identified a mean of 24.65 issues on the un-mutated contracts, with a median of 17 issues found; this was a significant difference in issue counts from either other tool, with $p<1.0^{-15}$ in both cases.  Despite this huge number of non-informational warnings for the original contracts, the mean mutation score was even worse than for SmartCheck, 0.03.  For contracts where Securify found no issues in the original contract, it \emph{never} detected any mutants. Securify, however, did detect 1,449 mutants not detected by Slither, 1,076 of which were not detected by SmartCheck.  On the other hand, given the ratio of issue counts compared to both tools is over 10.0, it is fairly likely many of these \emph{are} false positives, despite the probability mutants are actual faults.

Our recommendation, based on mutation analysis results, is that smart contract analysis should at least use both Slither and SmartCheck, perhaps applying only Slither until it reports no interesting issues.  Securify does offer a substantial non-overlapping set of issues reported, however, but produces a very large number of false positives.  Available human resources and total number of issues produced (as well as high runtime) might justify not using Securify in some cases.

\subsubsection{Improving Slither {\bf (RQ 4)}}

@@

\begin{figure*}
\begin{code}
  0x598ab825d607ace3b00d8714c0a141c7ae2e6822\_Vault.mutant.275.sol:
  424contracts/0x598ab825d607ace3b00d8714c0a141c7ae2e6822\_Vault.sol:249
        if (!p.recipient.send(p.amount)) \{  // Make the payment
 ==>          if (true) \{  // Make the payment
        if (true) \{  // Make the payment
      \end{code}
      \caption{Mutant showing {\bf Boolean constant misuse}.}
      \label{fig:boolean}
    \end{figure*}

    \begin{figure*}
\begin{code}
  0x968815CD73647C3af02a740a2438D6f8219e7534\_TTPresale.mutant.311.sol:
  424contracts/0x968815CD73647C3af02a740a2438D6f8219e7534\_TTPresale.sol:175
        require(nextDiscountTTMTokenId6 >= 361 \&\& nextDiscountTTMTokenId6 <= 391);
 ==>  ...361...==>...0...
        require(nextDiscountTTMTokenId6 >= 0 \&\& nextDiscountTTMTokenId6 <= 391);
      \end{code}
      \caption{Mutant showing {\bf Type-based tautologies}.}
      \label{fig:tautology}      
\end{figure*}

\begin{figure*}
\begin{code}
  0x534ccee849a688581d1b0c65e7ff317ed10c5ed3\_NametagToken.mutant.480.sol:
  424contracts/0x534ccee849a688581d1b0c65e7ff317ed10c5ed3\_NametagToken.sol:650
        byte char = byte(bytes32(uint(x) * 2 ** (8 * j)));
 ==>  ...*...==>.../...
        byte char = byte(bytes32(uint(x) * 2 ** (8 / j)));
      \end{code}
      \caption{Mutant showing {\bf Loss of precision due to divide before multiply}.}
      \label{fig:divmul}            
      \end{figure*}

@@

\item {\bf Loss of precision due to divide before multiply:}  Solidity only supports integer types.  This means that performing division before multiplication can introduce rounding that is not present when the multiplication is performed first.  This is a fairly important problem, given the frequency with which Solidity code performs financial calculations where maximum precision is desired.  SmartCheck provides a detector for such precision losses \url{https://github.com/smartdec/smartcheck/blob/master/rule\_descriptions/SOLIDITY\_DIV\_MUL/description\_en.html}, which enabled it to detect mutants such as the one shown in Figure\ref{fig:divmul}.
\end{enumerate}

@@

Examining the first 100 mutants in the unprioritized lists for SmartCheck and Securify, ordered by contract ID and mutant number (roughly source line mutated) we were unable to identify \emph{any} obviously interesting mutants.  The majority of cases involved either the missing return problem or replacing {\tt msg.sender} with {\tt tx.origin}; Slither has a detector for misuses of {\tt tx.origin}, and we believe (but are not sure) that almost all of these are due to intentional behavior: SmartCheck and Securify tend to identify most uses of {\tt tx.origin} as incorrect, while Slither has a more selective rule intended to avoid false positives.

It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying good candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ4}.

@@

@@

\begin{table}
  \begin{tabular}{l|r|r|r|r|r}
    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & Mutant \\
    Tool & Mean & Median & Mean & Median & Ratio\\
    \hline
    \hline
    SpotBugs & 0.26 & 0.00 & 0.30 & 0.07 & 1.132 \\
    \hline
    PMD & 0.48 & 0.00 & 0.09 & 0.00 & 0.183 \\
    \hline
    Infer & 0.00 & 0.00 & 0.00 & 0.00 &  0.006 \\
    \hline
  \end{tabular}
  \caption{Java tools: number of findings and mutation scores over all files.}
  \label{tab:scorejava}
\end{table}

\begin{table}
  \begin{tabular}{l|r|r|r|r|r}
    & & \multicolumn{2}{|c|}{} & \multicolumn{2}{|c}{Clean For All (1,185)} \\
    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} &  \multicolumn{2}{|c}{Mutation Score}\\
    Tool & Files & Mean & Median & Mean & Median\\
    \hline
    \hline
    SpotBugs & 1,426 & 0.32 & 0.05 & 0.37 & 0.06 \\
    \hline
    PMD & 1,359 & 0.08 & 0.00 & 0.08 & 0.00 \\
    \hline
    Infer & 1,598 & 0.00 & 0.00 & 0.00 & 0.00 \\
    \hline
  \end{tabular}
  \caption{Java tools: clean file counts and mutation scores.}
  \label{tab:cleanjava}
\end{table}

Figure \ref{fig:javavenn} shows the mutants killed by the Java analysis tools.

@@

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{python.png}
  \caption{Mutants killed by Python static analysis tools.}
  \label{fig:javavenn}
\end{figure}

\begin{table}
  \begin{tabular}{l|r|r|r|r|r}
    & \multicolumn{2}{|c|}{Findings} & \multicolumn{2}{|c|}{Mutation Score}  & Mutant \\
    Tool & Mean & Median & Mean & Median & Ratio\\
    \hline
    \hline
    SpotBugs & 0.26 & 0.00 & 0.30 & 0.07 & 1.132 \\
    \hline
    PMD & 0.48 & 0.00 & 0.09 & 0.00 & 0.183 \\
    \hline
    Infer & 0.00 & 0.00 & 0.00 & 0.00 &  0.006 \\
    \hline
  \end{tabular}
  \caption{Java tools: number of findings and mutation scores over all files.}
  \label{tab:scorejava}
\end{table}

\begin{table}
  \begin{tabular}{l|r|r|r|r|r}
    & & \multicolumn{2}{|c|}{} & \multicolumn{2}{|c}{Clean For All (1,185)} \\
    & \# Clean & \multicolumn{2}{|c|}{Mutation Score} &  \multicolumn{2}{|c}{Mutation Score}\\
    Tool & Files & Mean & Median & Mean & Median\\
    \hline
    \hline
    SpotBugs & 1,426 & 0.32 & 0.05 & 0.37 & 0.06 \\
    \hline
    PMD & 1,359 & 0.08 & 0.00 & 0.08 & 0.00 \\
    \hline
    Infer & 1,598 & 0.00 & 0.00 & 0.00 & 0.00 \\
    \hline
  \end{tabular}
  \caption{Java tools: clean file counts and mutation scores.}
  \label{tab:cleanjava}
\end{table}

@@
